\chapter{The System of Linear Transformations}
\label{chapter-linear-systems}
From now on, we concentrate on the systems of linear transformations between vector spaces over the field $\mathbb{Z}_2$. The reason is simple, they allow a construction of a remarkable model of universal hashing. Like for the other universal models, the expected length of a chain is constant and so is the expected running time of the find operation. In addition, the expected length of the longest chain is bounded by a sub-linear function of a single variable -- the size of the hash table. In our model, we have also managed to prove that the expected amortised time of every operation, find, insert and delete, is constant.

Once again, the main result is the upper bound on the expected length of the longest chain when a system of linear maps is used as a universal class. As already mentioned, probability of collision of $k$-elements is not estimated directly from the properties common to all universal systems. Hence the proof of the bound can not be based on the idea shown in the previous chapter.

First, we show a few models of the uniform random choice of a linear function. They describe other ways of the uniform random choice of a linear function from the universal class -- choice of the universal hash function. Later, we prove some technical lemmas regarding the vector spaces. Finally, we state the required facts in terms of system of linear transformations. Then we interpret the facts in terms of hashing and propose a new model. Many theorems come from \cite{DBLP:journals/jacm/AlonDMPT99}. Our work lies in their correction, improvement and modification leading to a better and practical result.

In the following sections we restrict ourselves only to the vector spaces over the field $\mathbb{Z}_2$. If vector spaces are taken over other finite field than the field $\mathbb{Z}_2$, then we can not expect such good results as shown in \cite{DBLP:journals/jacm/AlonDMPT99}. The original result proposes hashing of $m \log m$ elements into a hash table of size of $m$ slots and leads to the bound $O(\log m \log \log m)$. Although the result may already be used for hashing $m$ elements, too, it is later improved for $\alpha \leq 1$.

When unsure about any exact definition or fact from linear algebra, see Appendix~\ref{appendix-linear-algebra} where we exactly summarise and accurately state them.

\input{linear_transformations_models.tex}
\input{linear_transformations_probability.tex}
\input{linear_transformations_distribution.tex}
\input{linear_transformations_elpsl.tex}
\input{linear_transformations_constant_estimate.tex}
