\documentclass{article}
\usepackage{wds10,epsf}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amsthm}
\include{definitions}
\usepackage{graphicx}
\usepackage{url}
\usepackage{algpseudocode}
\usepackage{algorithm}

\lefthead{Martin Babka}
\righthead{A Worst Case Aware Version of Universal Hashing}

\setcounter{secnumdepth}{0}

\begin{document}
\title{A Worst Case Aware Version of Universal Hashing}

\author{Martin Babka}

\affil{Charles University, Faculty  of  Mathematics  and  Physics,
     Prague, Czech Republic.}

\begin{abstract}
In this work we propose and analyze a simple hash table giving a worst case time guarantee in case of Find operation with all operations running in constant expected amortized time.
To achieve this goal we exploit known properties of some universal classes of functions.
Since the mentioned guarantee strongly depends on the quality of the underlying universal system, we show various systems capable of interesting bounds.
With reasonable assumptions we can also combine the proposed model with the "two choices paradigm" and obtain $\bigo(\log \log n)$ worst case bound.
The main advantage of the method is its simplicity and genericness because it may be used with separating chaining as well as with linear probing.
\end{abstract}

\begin{article}

\input{introduction}
\input{model}
\input{limit_function}
\input{conclusion}

% TODO: Cislo GAUK-u
\acknowledgments %\footnotesize
{We would like to thank Václav Koubek for helpful comments.
The  present  work  was  supported by
the Czech  Grant  Agency  under  Contracts  205/99/1712, and
205/00/1686 and by the Charles University Grant Agency under
Contract 181.}

\input{bibliography}

\end{article}
\end{document}