\section{Conclusion}
\label{section-conclusion}
We proposed a simple modification of universal hashing which guarantees worst case time for a look up. Our approach works well not only with separate chaining but with linear probing or double hashing as well if the distribution or at least the expected value of length of the longest probe sequence is known. Malalla showed in his dissertation \cite{Malalla:2004:THS:1124034} how to use the two choices paradigm with linear probing in case of plain hashing. Now we have two possibilities how to apply the result. If the input data is random enough, we can assume full randomness of the hash function. Naturally, we can use an almost uniform system in the general case. This solution achieves a limit function $\bigo(\log \log n)$ on the length of a probe sequence.

Another solution, which works without any further assumptions on the input data and without uniform systems, is to study behavior of universal hashing with efficient and natural classes. An interesting problem with the worst case aware hashing is a way how to derive bounds for two choices paradigm connected with a reasonable universal system. Our simulations show that using simple systems, e.g. functions of the form $((ax + b) \bmod |U|) \bmod |V|$ or higher degree polynomials for $|U|$ being a prime are not impressive. Hashing $2^{29}$ elements with these universal systems results in longest chains containing about $250$ elements for certain types of sets. However, with system of all linear transformations we were not able to find a set having more than $15$ elements for the same size of the stored set.

As showed in \cite{DBLP:conf/alenex/ThorupZ10} linear probing is also possible with universal hashing provided that at least $5$-wise independent family of functions is used. However, bounds on the length of the longest chain are not known so far. Because of the cache behavior currently fastest hash tables are based on linear probing. Therefore obtaining bounds on $\lpsl$ for universal hashing is an interesting problem. It could bring a fast hash table giving warranties for the worst case look up time. In addition two choices paradigm may be helpful to significantly improve the warranty.