% This is the LaTeX source for the article dealing with Worst Case 
% Aware Universal Hashing. The source code is a modification of the 
% file typeinst.tex which may be found at http://www.springer.com/lncs
% and ftp://ftp.springer.de/data/pubftp/pub/tex/latex/llncs/latex2e

\documentclass[runningheads,a4paper]{llncs}
\include{definitions}

\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\setcounter{tocdepth}{3}
\usepackage{graphicx}
\usepackage{url}

\urldef{\mail}\path|babkys@gmail.com,vcunat@matfyz.cz|
\newcommand{\keywords}[1]{\par\addvspace\baselineskip
\noindent\keywordname\enspace\ignorespaces#1}
\usepackage[utf8]{inputenc}

\begin{document}

\mainmatter

% TODO: Coslo GAUKu dopisat.
\author{Martin Babka\thanks{Moj GAUK} \and Vladimír Čunát\thanks{This research was partially supported by SVV project number 263 314.}}
\authorrunning{Martin Babka, Vladimír Čunát}

\title{A Worst Case Aware Version of Universal Hashing}
\titlerunning{A Worst Case Aware Version of Universal Hashing}
\toctitle{A Worst Case Aware Version of Universal Hashing}

\institute{Faculty of Mathematics and Physics, Charles University,\\ Prague, Czech Republic \\
\mail\\
\url{http://www.ktiml.mff.cuni.cz/~babka/hashing}}

\tocauthor{Martin Babka, Vladimír Čunát}
\maketitle

\begin{abstract}
In this work we propose and analyze a simple hash table giving a worst case time guarantee in case of Find operation with all operations running in constant expected amortized time.
To achieve this goal we exploit known properties of some universal families of functions.
Since the mentioned guarantee strongly depends on the quality of the underlying universal system, we show various systems capable of interesting bounds.
With reasonable assumptions we can also combine the proposed model with the "two choices paradigm" and obtain $\bigo(\log \log n)$ worst case bound.
The main advantage of the method is its simplicity and genericness because it may be used with separating chaining as well as with linear probing.

\keywords{universal hashing, data structures, worst case lookup}
\end{abstract}

\input{introduction}
\input{model}
\input{limit_function}
\input{conclusion}

\subsubsection*{Acknowledgments.}
% TODO: Lepsia formulacia (others).
We would like to thank Václav Koubek and some others for their helpful comments.

\input{bibliography}

\clearpage
% \input{appendix_amortised_expected_time}
\input{appendix_linear_functions}
\input{appendix_two_choices_uniform}

\end{document}
